\documentclass[letterpaper,10pt,titlepage]{article}

\setlength{\parindent}{0pt}

\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}

\usepackage{alltt}
\usepackage{float}
\usepackage{color}
\usepackage{url}
\usepackage{listings}

\usepackage{balance}
\usepackage[TABBOTCAP, tight]{subfigure}
\usepackage{enumitem}
\usepackage{pstricks, pst-node}

\usepackage{geometry}
\geometry{textheight=8.5in, textwidth=6in}

\newcommand{\cred}[1]{{\color{red}#1}}
\newcommand{\cblue}[1]{{\color{blue}#1}}

\usepackage{hyperref}
\usepackage{geometry}
\usepackage{vhistory}

\hypersetup{%
	colorlinks = true,
	linkcolor = black
}

\lstdefinestyle{customc}{
  belowcaptionskip=1\baselineskip,
  breaklines=true,
  frame=L,
  xleftmargin=\parindent,
  language=C,
  showstringspaces=false,
  basicstyle=\footnotesize\ttfamily,
  keywordstyle=\bfseries\color{green!40!black},
  commentstyle=\itshape\color{purple!40!black},
  identifierstyle=\color{blue},
  stringstyle=\color{orange},
}

\def\name{Alex Hoffer}

%pull in the necessary preamble matter for pygments output
\input{pygments.tex}

%% The following metadata will show up in the PDF properties
\hypersetup{
  colorlinks = true,
  urlcolor = black,
  pdfauthor = {\name},
  pdfkeywords = {CS444 ``Operating Systems''},
  pdftitle = {CS 444 Writeup 1},
  pdfsubject = {CS 444 Writeup 1},
  pdfpagemode = UseNone
}

\begin{document}

\begin{titlepage}
    \begin{center}
        \vspace*{3.5cm}

        \textbf{Writeup 1}

        \vspace{0.5cm}

        \textbf{Alex Hoffer\\}
	\textbf{Nehemiah Edwards\\}
	\textbf{Ryan Kitchen}

        \vspace{0.8cm}

        CS 444\\
        Spring 2017\\

        \vspace{1cm}

        \textbf{Abstract}\\

        \vspace{0.5cm}

        This first write up describes important details from the completion of Project 1 for D. Kevin McGrath's Operating Systems II class. Topics from Project 1 that are to be covered in this work include the building of the Linux Yocto kernel on Oregon State's engineering server, usage of the qemu virtual machine, and a solution of the Producer-Consumer concurrency problem using the C programming language's POSIX threads execution model. 

        \vfill

    \end{center}
\end{titlepage}

\newpage

\tableofcontents

\newpage

\section{Log of Commands to Build Yocto Kernel and Load Qemu}
\begin{lstlisting}
cd /scratch/spring2017/
mkdir 10-04
git clone git://git.yoctoproject.org/linux-yocto-3.14
cd linux-yocto-3.14
git checkout v3.14.26
source /scratch/opt/environment-setup-i586-poky-linux
cp /scratch/spring2017/files/config-3.14.26-yocto-qemu .config
make menuconfig
make -j4 all
cd ..
gdb

(terminal #2)
source /scratch/opt/environment-setup-i586-poky-linux
cd /scratch/spring2017/10-04
cp /scratch/spring2017/files/core-image-lsb-sdk-qemux86.ext3 .
/scratch/spring2017/files/core-image-lsb-sdk-qemux86.ext3 .
qemu-system-i386 -gdb tcp::6504 -S -nographic -kernel bzImage-qemux86.bin -drive file=core-image-lsb-sdk-qemux86.ext3,if=virtio -enable-kvm -net none -usb -localtime --no-reboot --append "root=/dev/vda rw console=ttyS0 debug

(terminal 1)
gdb
target remote :6504
continue

(terminal 2)
root
uname -a
reboot
qemu-system-i386 -gdb tcp::6504 -S -nographic -kernel linux-yocto-3.14/arch/x86/boot/bzImage  -drive file=core-image-lsb-sdk-qemux86.ext3,if=virtio -enable-kvm -net none -usb -localtime --no-reboot --append "root=/dev/vda rw console=ttyS0 debug

(terminal 1)
gdb
target remote :6504
continue

(terminal 2)
root
uname -a
reboot
q
\end{lstlisting}

\section{Writeup of Concurrency Solution}
To solve the Producer-Consumer problem, I used POSIX threads and the -lpthreads flag on the command line to compile the program. Specifically, I made functions that corresponded to producing an item and consuming an item, I created a producer thread and passed it the producing an item void function, I created a consumer thread and passed it the consuming an item void function, and then I ran a loop joining the consumer thread to the producer thread. Within both producer/consumer functions, mutexes were used to lock and then unlock resources that could be subject to overwriting and ASM was used to randomly generate fields for item structures as well as waiting times for producer threads. A producer thread blocked, or "waited" for a consumer thread to consume an item, while a consumer thread "waited" if there was nothing to consume.

\section{Flags in the listed Qemu command line}
The listed Qemu command line is:
\begin{lstlisting}
qemu-system-i386 -gdb tcp::???? -S -nographic -kernel bzImage-qemux86.bin 
\end{lstlisting}
\begin{lstlisting}
-drive file=core-image-lsb-sdk-qemux86.ext3,if=virtio -enable-kvm 
\end{lstlisting}
\begin{lstlisting}
-net none -usb -localtime --no-reboot --append 
\end{lstlisting}
\begin{lstlisting}
"root=/dev/vda rw console=ttyS0 debug".
\end{lstlisting}

The following list describes each flag:
\begin{itemize}
\item \emph{-gdb} tells the system to wait to connect to gdb on the device tcp::????, this device being passed in as an argument.
\item \emph{-S} is a flag that tells the system to wait for a gdb connection.
\item \emph{-nographic} is a visualization option that turns off graphical output.
\item \emph{-kernel} is a flag that saves time and resources by granting the system a Linux kernel image, rather than requiring a full boot.
\item \emph{-drive} handles disk options and takes 1 to n number of arguments to specify these options, in the case of this command the file we passed to this flag is parsed according to the virtio interface.
\item \emph{-enable-kvm} is a flag that enables KVM virtualization support without any setting turned off.
\item \emph{-net} is a flag that sets network options.
\item \emph{-usb} is a flag that turns on the USB driver.
\item \emph{-localtime} is a flag that tells the system to use local time.
\item \emph{--no-reboot} tells the system to exit, not reboot, when rebooting is an option.
\item \emph{--append} is a flag that represents the command line necessary to interact with the kernel.
\end{itemize}

\section{Concurrency}
The following subsections answer the four questions as outlined on the Project 1 page on Kevin McGrath's course website.
\subsection{Main point of assignment}
There were several reasons for the assignment. One, I think, was to re-familiarize us with POSIX Threads, which we learned about in CS344 but likely forgot about before we got to this class. Second, I think that it was to teach us about the Producer-Consumer problem, which is a canonical example of concurrency. Third, it was to force us to use inline assembly language in our programs to help us learn how to implement ASM in C.
\subsection{Personal approach to problem}
My approach for the use of threads came from guidelines to the canonical Producer-Consumer problem that I found in the \emph{Linux Programming Interface} text. For example, statements like pthread\_create, pthread\_join, pthread\_t, pthread\_mutex\_t, and pthread\_cond\_t were all found in that book and that book instructed me on how to use them. I used these static pthread statements rather than dynamic ones because frankly, I could not think of a reason why dynamic ones would be necessary for the purposes of this assignment. My approach for the buffer and the item to be held in the buffer was just to use a struct to represent the buffer and a struct to represent the item and have the buffer hold a statically allocated array of pointers to items and for each item to have two data fields: a random number and a random waiting period, both that had to be implemented in ASM. Given the statements necessary to control threads and mutexes and the structs necessary to store data, I solved the problem by creating a thread to correspond to a Producer, having that Producer thread lock all of the shared data using a mutex and then produce a random item. After unlocking the shared data within the Producer thread, I created a Consumer thread that locked the data again, consumed the first non-null index in the buffer, and then unlocked the data again. I contained these Producer and Consumer thread creation statements in a for loop that runs the number of times the user indicates in a command line argument so that when the program is run it creates the corresponding producer and consumer threads and produces and consumes random data. 
\subsection{Ensuring solution was correct}
To ensure that the buffer was properly being added/removed to and to ensure that both threads were properly locking and unlocking, I added print statements in important functions such as when the global mutex was locked and unlocked in order to know whether these calls were actually being made. It is easy to verify that the threads are switching back and forth because both threads have different print statements that are only accessed within their specific thread is running. It is also easy to realize that the buffer is only sequentially being accessed by the two different threads in accordance with the mutex locking and unlocking because otherwise both threads would be accessing the same resource and either a memory read/write issue would occur or both threads would be depending on each other for an inordinate amount of time, resulting in \emph{deadlock}. Obviously, my use of the pthread condition functionality allowed both threads to wait until resources were available, thus preventing deadlock. Finally, to ensure that the entire integrated software worked properly, I used a makefile to generate an executable of the code that accepts 1 argument that corresponds to the number of times a loop that creates the threads shall occur. To ensure the executable achieved what I wanted it to, I ran the software with command line arguments of 1, 2, 3, 4, and 5. The corresponding output for these arguments properly demonstrated that the different threads were called 1, 2, 3, 4, and 5 times, respectively, which is correct behavior.

\subsection{What I learned}
I learned more about POSIX thread functions, such as initializing them, the act of "joining" them, and locking resources down using a mutex to prevent threads from accessing the same data simultaneously. I also learned more about implementing in-line ASM into the C programming language, a skill that could pose useful for software projects that require maximized efficiency in the context of low-level programming. I found the in-line ASM to be complex and confusing, mainly because of a) the odd syntax and b) how it required my brain to be thinking in both ASM and C at the same time.

\section{Version control log}
\begin{versionhistory}
\vhEntry{1}{4.7.2017}{Alex}{Add latex template file}
\vhEntry{2}{4.9.2017}{Alex}{Created structs to hold the data, inspect how to do the ASM randomization}
\vhEntry{3}{4.9.2017}{Alex}{Add fake random functions to stand in place until ASM is understood}
\vhEntry{4}{4.9.2017}{Alex}{add pseudocode for producer, consumer void functionswhich are passed to pthreads}
\vhEntry{5}{4.10.2017}{Alex}{First pass at producer function done with mutexes, conditions, etc.}
\vhEntry{6}{4.10.2017}{Alex}{Everything done except ASM randomization, currently prog loops forever, only terminating when hit with a signal}
\vhEntry{7}{4.17.2017}{Alex}{change name of proj1 folder to project1}
\vhEntry{8}{4.17.2017}{Alex}{Add part in latex about Qemu command line, flags explained}
\vhEntry{9}{4.17.2017}{Alex}{Answer first two questions of the concurrency writeup}
\vhEntry{10}{4.18.2017}{Alex}{Add loop counter accessible from command line}
\vhEntry{11}{4.18.2017}{Alex}{Changed all typedefs of structs to just structs in accordance with Linus's coding style}
\vhEntry{12}{4.18.2017}{Alex}{Add what I learned and how I ensured solution was correct sections to writeup}
\vhEntry{13}{4.18.2017}{Alex}{Begin work on version control table}
\vhEntry{14}{4.19.2017}{Alex}{implementing ASM randomizer}
\vhEntry{15}{4.19.2017}{Alex}{Finish using ASM for randomization}
\vhEntry{16}{4.19.2017}{Alex}{add nehemiah's latex part}
\end{versionhistory}

\section{Work log}
\subsection{April 7-9}
Alex added the Latex template file and got accustomed to using Kevin's makefile utilizing the commands \emph{latex latex.tex} and \emph{make}. He created the structs to hold the data as the assignment specified and looked into doing the ASM randomization. He made fake random functions in lieu of using the ASM which will be implemented last so he could begin primary development.
\subsection{April 10-17}
Alex used his Linux programming book to re-familiarize himself with mutexes, pthreads, etc. He was delighted to find examples of the Producer-Consumer problem in his book. He took a first pass at both void Producer-Consumer functions, complete with mutexes. He was also delighted to find out that both functions were easy to implement. Then, he began to do the Latex writeup. He wrote the abstract and got the basic structure of the file's sections and subsections established. Then, he wrote the part about the Qemu command line where he explained each flag that was used. Then, he answered the first two questions of the concurrency writeup section.  
\subsection{April 18-19}
Alex read the Coding Style requirements on Kevin's page and got rid of the typedef'd structs he had set up initially because these were not kosher. He added a loop counter accessible from the command line so that for testing the software he could run the thread loop a set number of times. Then, he added the "What I learned" and "How I ensured the solution was correct" sections to the writeup, and then figured out how to make a nice looking version control log. For the finale of the project, Nehemiah wrote the list of all commands necessary for kernel build/yocto and Alex implemented ASM randomization functions using several different online resources (these resources are indicated in the comments of the code). 

%\bibliographystyle{plain}
%\bibliography{CS444_Writeup1}
\end{document}
